3. Analyses Plan

### 2021-06-21 reports/01_replication.Rmd
prepare pre-registration

library(haven)
library(data.table)
library(magrittr)
library(kableExtra)
library(DescTools) # gini coefficient
library(downloadthis)
library(Rmisc)
library(ggplot2)
library(wesanderson)

DT <- read_dta(file="../data/gaechteretal/GMTV-data.dta") %>% data.table()
```

@@

```{r data, include = FALSE}
noPunish10 <- DT[exp_num == 5 | exp_num == 8 | exp_num == 9]
```

@@

```{r writeData}

noPunish10[,
           share := sum/gdp,
           by = .(subj_id, per)]

gaechter <- noPunish10[order(gr_id, per),
           .(treatment = "noPunish10", 
             session.code = exp_num,
             groupID = gr_id, 
             round = per,
             contribution = sum,
             endowment = gdp,
             share = mean, # share,
             stock = (gdp + ceiling(sum/4*1.5)*4-sum),
             gain = (ceiling(sum/4*1.5)*4-sum),
             gini = gini2,
             bot_active = 0,
             # disaster = 0,
             # EWE = 0,
             noComprehension = 0)] %>% unique()

gaechter %>%
  head(n = 15) %>%
  kbl() %>% 
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

save(gaechter, file = "../data/processed/rda/gaechter2017.rda")
write.csv(gaechter, file ="../data/processed/csv/gaechter2017.csv")

@@

```{r downloadData}
gaechter %>%
      download_this(
        output_name = "Gaechter2017noPunish10",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

```{r plotContribution}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "contribution", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotShare}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "share", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotStock}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "stock", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotGini}
plotDT <- summarySE(data = gaechter, 
                      measurevar = "gini", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

gaechterFirstRound <- temp[,
                           .(participant.code = subj_id,
                             treatment,
                             session.code = exp_num,
                             groupID = gr_id,
                             othersContribution = sumputin - putin,
                             ownContribution = putin,
                             trust = NA,
                             comprehension = NA)]

@@

# save data
fileName <- "gaechter2017"
save(gaechterFirstRound, 
     file = paste0("../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(gaechterFirstRound, 
          file = paste0("../data/processed/csv/", fileName, "_R1", ".csv"))
```

`r NROW(gaechterFirstRound)` participants yield `r NROW(gaechterFirstRound)` independent observations that are stored in a data table called ```gaechterFirstRound```. These observations are displayed below and can, once more, be downloaded by a click on the button. As before, the data is also saved in ```../data/processed/```.

```{r displayFirstRound}
# display data
gaechterFirstRound %>%
  head(n = 15) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

# create download button
gaechterFirstRound %>%
      download_this(
        output_name = "gaechter2017_R1",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

>>>>

library(haven)
library(data.table)
library(magrittr)
library(kableExtra)
library(DescTools) # gini coefficient
library(downloadthis)
library(Rmisc)
library(ggplot2)
library(wesanderson)
library(lubridate)
```

```{r readData}
DT <- read_dta(file="../data/gaechteretal/GMTV-data.dta") %>% data.table()
CT <- read_dta(file="../data/gaechteretal/GMTV-questionnaire-data.dta") %>% data.table()
```

@@

```{r data, include = FALSE}
noPunish10 <- DT[exp_num == 5 | exp_num == 8 | exp_num == 9]
# noPunish10 <- DT[longgame == 0 & punish == 0 & exp_num <= 10]
```

@@

```{r manipulateData}
noPunish10[,
           share := sum/gdp,
           by = .(subj_id, per)]

GMTV <- noPunish10[order(gr_id, per),
           .(treatment = "noPunish10", 
             session.code = exp_num,
             groupID = gr_id, 
             round = per,
             contribution = sum,
             endowment = gdp,
             share = mean, # share,
             stock = (gdp + ceiling(sum/4*1.5)*4-sum),
             gain = (ceiling(sum/4*1.5)*4-sum),
             gini = gini2,
             bot_active = 0,
             # disaster = 0,
             # EWE = 0,
             noComprehension = 0)] %>% unique()

# create rich indicator (just as highgdp in original data)
median <- GMTV[round == 10,
                   median(stock)]

richGroups <- GMTV[round == 10 & stock > median,
                       unique(groupID)]

poorGroups <- GMTV[round == 10 & stock < median,
                       unique(groupID)]

GMTV[groupID %in% richGroups,
         rich := TRUE]

GMTV[groupID %in% poorGroups,
         rich := FALSE]

# display data
GMTV %>%
  head(n = 15) %>%
  kbl() %>% 
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

@@

```{r writeData}
save(GMTV, file = "../data/processed/rda/GMTV2017.rda")
write.csv(GMTV, file ="../data/processed/csv/GMTV2017.csv")

```{r downloadData}
GMTV %>%
      download_this(
        output_name = "GMTV2017noPunish10",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

```{r plotContribution}
plotDT <- summarySE(data = GMTV, 
                      measurevar = "contribution", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotShare}
plotDT <- summarySE(data = GMTV, 
                      measurevar = "share", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotStock}
plotDT <- summarySE(data = GMTV, 
                      measurevar = "stock", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

```{r plotGini}
plotDT <- summarySE(data = GMTV, 
                      measurevar = "gini", 
                      groupvars=c("treatment", "round"),
                      na.rm = FALSE,
                      conf.interval = 0.95,
                      .drop = TRUE)

@@

GMTVFirstRound <- temp[,
                           .(participant.code = subj_id,
                             treatment,
                             session.code = exp_num,
                             groupID = gr_id,
                             othersContribution = sumputin - putin,
                             ownContribution = putin,
                             trust = NA,
                             comprehension = NA)]

@@

# save data
fileName <- "GMTV2017"
save(GMTVFirstRound, 
     file = paste0("../data/processed/rda/", fileName, "_R1", ".rda"))
write.csv(GMTVFirstRound, 
          file = paste0("../data/processed/csv/", fileName, "_R1", ".csv"))
```

`r NROW(GMTVFirstRound)` participants yield `r NROW(GMTVFirstRound)` independent observations that are stored in a data table called ```GMTVFirstRound```. These observations are displayed below and can, once more, be downloaded by a click on the button. As before, the data is also saved in ```../data/processed/```.

```{r displayFirstRound}
# display data
GMTVFirstRound %>%
  head(n = 15) %>%
  kbl() %>% 
  # scroll_box(height = "200px") %>%
  kable_paper("hover", 
              full_width = TRUE,
              fixed_thead = TRUE)

# create download button
GMTVFirstRound %>%
      download_this(
        output_name = "GMTV2017_R1",
        output_extension = ".csv", # CSV output
        button_label = "Download csv",
        button_type = "default"
        )
```

@@

```{r subsetCovs}
GMTVCovariates <- CT[exp_num == 5 | exp_num == 8 | exp_num == 9,
                         .(treatment = "noPunish10",
                           session.code = exp_num,
                           groupID = gr_id,
                           participant.code = subj_id,
                           gender,
                           age = Sys.Date() %>% lubridate::year() - age,
                           pq01 = q1,
                           pq02 = q2,
                           pq03 = q3,
                           pq04 = q4,
                           pq05 = q5,
                           pq06 = q6,
                           pq07 = q7,
                           pq08 = q8,
                           pq09 = q9,
                           pq10 = q10,
                           pq11 = q11,
                           pq12 = q12,
                           pq13 = q13,
                           pq14 = q14,
                           donation = don
                           )]

# write data
fileName <- "GMTV2017"
save(GMTVCovariates, 
     file = paste0("../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(GMTVCovariates, 
          file = paste0("../data/processed/csv/", fileName, "_COVS", ".csv"))

# no display or download button as these data were not made available online by the authors
```

### 2021-06-21 reports/02_dataProcessing.Rmd
prepare pre-registration

```{r loadData}
# read simulated data
DT <- read.csv("../data/simulation/all_apps_wide.csv", stringsAsFactors = FALSE) %>%
  data.table()
```

@@

Eventually, this report can also be used to process real data. This only requires us to load a different data set.^[It may also require us to merge multiple data sets -- that depends on whether the lab delivers one or several data sets and remains to be discussed.]

@@

```{r firstRound}
# create data table 
replicationFirstRound <- DT[,
                 .(participant.code,
                   treatment = "replication",
                   session.code,
                   groupID = dPGG.1.group.id_in_subsession,
                   othersContribution = dPGG.1.group.total_contribution - dPGG.1.player.contribution,
                   ownContribution = dPGG.1.player.contribution,
                   trust = Outro.1.player.PQ11,
                   comprehension = dPGG.10.player.comprehension)]

@@

@@

>>>>

```{r readData}
# read simulated data
DT <- read.csv("../data/simulation/all_apps_wide.csv", stringsAsFactors = FALSE) %>%
  data.table()
```

@@

Eventually, this report can also be used to process real data. This only requires us to read a different data set.^[It may also require us to merge multiple data sets -- that depends on whether the lab delivers one or several data sets and remains to be discussed.]

@@

```{r firstRound}
# create data table 
replicationFirstRound <- DT[,
                 .(participant.code,
                   treatment = "replication",
                   session.code,
                   groupID = paste(session.code, dPGG.1.group.id_in_subsession, sep = "_"),
                   othersContribution = dPGG.1.group.total_contribution - dPGG.1.player.contribution,
                   ownContribution = dPGG.1.player.contribution,
                   trust = Outro.1.player.PQ11,
                   comprehension = dPGG.10.player.comprehension)]

@@

Next, we'll flag groups that are richer and poorer than the median group.

```{r calcRich}

median <- replication[round == 10,
                      median(stock)]

richGroups <- replication[round == 10 & stock > median,
                          unique(groupID)]

poorGroups <- replication[round == 10 & stock < median,
                          unique(groupID)]

replication[groupID %in% richGroups,
            rich := TRUE]

replication[groupID %in% poorGroups,
            rich := FALSE]

```

@@

```{r createCovs}

# add variables
DT[, treatment := "replication"]

DT[, groupID := paste(session.code, dPGG.1.group.id_in_subsession, 
                    sep = "_")]

# subset
cRegex <- "participant.code|session.code|treatment|groupID|Outro.1.player|10.player.donation"
covariates <- str_subset(string = names(DT),
                            pattern = cRegex)
CT <- DT[, ..covariates]

# rename
names(CT) <- names(CT) %>% 
  str_replace_all(pattern =".*player\\.",
                  replacement = "") %>%
  str_to_lower()

names(CT)[names(CT) == "groupid"] <- "groupID"

# refactor
CT[, donation := donation/20]

CT[, gender := ifelse(test = gender == "female",
                      yes  = 1,
                      no   = 0)]

replicationCovariates <- CT

# write data
fileName <- "replication2021"
save(replicationCovariates, 
     file = paste0("../data/processed/rda/", fileName, "_COVS", ".rda"))
write.csv(replicationCovariates, 
          file = paste0("../data/processed/csv/", fileName, "_COVS", ".csv"))

# no display or download button as these data may qualify as PII
```

### 2021-06-21 reports/03_dataAnalyses.Rmd
prepare pre-registration

library(data.table)
library(magrittr)
library(kableExtra)
library(downloadthis)
library(ggplot2)
library(wesanderson)
```

@@

```{r readData}
load(file="../data/processed/rda/gaechter2017.rda")        # read gaechter
load(file="../data/processed/rda/gaechter2017_R1.rda")     # read gaechterFirstRound
load(file="../data/processed/rda/replication2021.rda")     # read replication
load(file="../data/processed/rda/replication2021_R1.rda")  # read replicationFirstRound
```

```{r cbind}
main <- rbindlist(list(replication, gaechter), 
                  use.names = TRUE)

R1   <- rbindlist(list(replicationFirstRound, gaechterFirstRound), 
                  use.names = TRUE)
```

# Graphs

@@

```{r firstRound}
ggplot(R1, aes(y=othersContribution, x=belief)) +
  layout +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", alpha = 0.66) +
  geom_point() +
  geom_point(color = wes_palette("Moonrise3")[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 60)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 60))

@@

>>>>

library(data.table)
library(magrittr)
library(kableExtra)
library(downloadthis)
library(ggplot2)
library(wesanderson)
library(patchwork)
library(stargazer)
```

@@

```{r readData}
load(file="../data/processed/rda/GMTV2017.rda")          # read GMTV
load(file="../data/processed/rda/GMTV2017_R1.rda")       # read GMTVFirstRound
load(file="../data/processed/rda/GMTV2017_COVS.rda")     # read GMTVCovariates
load(file="../data/processed/rda/replication2021.rda")       # read replication
load(file="../data/processed/rda/replication2021_R1.rda")    # read replicationFirstRound
load(file="../data/processed/rda/replication2021_COVS.rda")  # read replicationCovariates
```

```{r cbind}
main <- rbindlist(list(replication, GMTV), 
                  use.names = TRUE)

R1   <- rbindlist(list(replicationFirstRound, GMTVFirstRound), 
                  use.names = TRUE)

covs <- rbindlist(list(replicationCovariates, GMTVCovariates), 
                  use.names = TRUE,
                  fill = TRUE)

rm(list = c("replication", "GMTV", "replicationFirstRound", "GMTVFirstRound", "GMTVCovariates", "replicationCovariates"))
```

# Results

@@

## First Round

```{r firstRound, eval = FALSE}
ggplot(R1, aes(y=othersContribution, x=belief)) +
  layout +
  geom_abline(intercept = 0, slope = 1, linetype="dashed", alpha = 0.66) +
  geom_point() +
  geom_point(color = wes_palette("Moonrise3")[1]) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 60)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 60))

@@

## Provision of the public good and wealth creation

```{r plotContributions}

SUM <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "contribution"]

SUM[,
    sum := round(contribution/4)]

upperLimit <- SUM$contribution %>% max() %>% round() + 10

p1 <- ggplot(data = SUM, 
             aes(x = round, y = contribution, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Average Amount of Tokens contributed") +
  scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("SUM"))
```

```{r plotShareOfContributions, fig.cap = "The average amount of tokens contributed over time in treatments."}

SHARE <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "share"]

upperLimit <- 1

p2 <- ggplot(data = SHARE, 
             aes(x = round, y = share, fill = treatment, color = treatment, lty = treatment)) +
  layout +
  theme(legend.position="bottom") +
  geom_line(show.legend=FALSE) +
  geom_point() +
  scale_x_continuous(name="",  breaks = 1:15) +
  scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
  labs(y = "Share of Current Endowment contributed") +
  scale_color_manual(values = wes_palette("GrandBudapest1"))

patchwork <- p1 + p2 & theme(legend.position = "bottom")
patchwork <- patchwork + plot_layout(guides = "collect")
patchwork

rm(list = c("SHARE", "p1", "p2"))
```

```{r summary1, results = "asis", warning = FALSE, fig.cap = "test"}

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

replication <- main[treatment == "replication" & round == 10, stock]
GMTV    <- main[treatment == "noPunish10" & round == 10, stock]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

rs1 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

The rank sum test yields a p-Value of `r rs1` for the mean ```stock``` a.k.a _Wealth_ during the last period of the game.

```{r tableStock, results = "asis", warning = FALSE}

# create subsets
main_all  <- main[round == 10]
main_poor <- main[round == 10 & rich == FALSE]
main_rich <- main[round == 10 & rich == TRUE]

# create table
if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

stargazer(lm(formula = stock ~ treatment, data = main_all),
          lm(formula = stock ~ treatment, data = main_poor),
          lm(formula = stock ~ treatment, data = main_rich),

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Wealth (Table 2)",
          header=FALSE,
          covariate.labels = c("Replication"),

          type = type, digits = 2, omit.stat = c("adj.rsq", "f"), df = FALSE
          )
```

```{r plotStock, fig.cap = "Average wealth over time across treatments."}
STOCK <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "stock"]

upperLimit <- STOCK$stock %>% max() %>% round() + 10

ggplot(data = STOCK, 
       aes(x = round, y = stock, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Wealth") +
          scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("STOCK"))
```

## Inequality

```{r plotInequality, fig.cap = "Average Gini coefficient over time across treatments."}
GINI <- main[,
            lapply(.SD, mean, na.rm = TRUE),
            by = c("round", "treatment"),
            .SDcols = "gini"]

upperLimit <- GINI$gini %>% max() %>% round(digits = 1) + 0.10

ggplot(data = GINI, 
       aes(x = round, y = gini, fill = treatment, color = treatment, lty = treatment)) +
          layout +
          theme(legend.position="bottom") +
          # geom_vline(xintercept = 10, alpha = 0.66) +
          geom_line(show.legend=FALSE) +
          geom_point() +
          scale_x_continuous(name="",  breaks = 1:10) +
          scale_y_continuous(limits = c(0, upperLimit), expand = c(0, 0)) +
          labs(y = "Gini coefficient") +
          scale_color_manual(values = wes_palette("GrandBudapest1"))

rm(list = c("GINI"))
```

```{r tableGINI, results = "asis", warning = FALSE}

# create subsets
main_all  <- main[round == 10]
main_poor <- main[round == 10 & rich == FALSE]
main_rich <- main[round == 10 & rich == TRUE]

# create table
if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

stargazer(lm(formula = gini ~ treatment, data = main_all),
          lm(formula = gini ~ treatment, data = main_poor),
          lm(formula = gini ~ treatment, data = main_rich),

          column.labels = c("All", "Below median", "Above median"),
          model.numbers = FALSE,
          dep.var.labels = "Wealth (Table 2)",
          header=FALSE,
          covariate.labels = c("Replication"),

          type = type, digits = 2, omit.stat = c("adj.rsq", "f"), df = FALSE
          )
```

```{r summary2, results = "asis", warning = FALSE}

if(knitr::is_html_output()){
  type <- "html"
} else {
  type = "latex"
}

replication <- main[treatment == "replication" & round == 10, gini]
GMTV    <- main[treatment == "noPunish10" & round == 10, gini]

rows <- sapply(X = list(replication, GMTV), FUN = NROW) %>% max()
temp <- data.frame(replication = c(replication, rep(NA, rows - NROW(replication))),
                   GMTV = c(GMTV, rep(NA, rows - NROW(GMTV))))

stargazer(temp,
          summary.stat = c("mean", "median","sd", "max", "min", "n"),
          type = type, 
          flip = TRUE, 
          header=FALSE)

rs2 <- wilcox.test(replication, GMTV)$p.value %>% round(digits = 4) %>% formatC(format = "f", digits = 4)

The rank sum test yields a p-Value of `r rs2` for the mean ```gini``` during the last period of the game.

## Sample Properties
